First look at Spacetime

### 2019-10-28 Output/error_reduction_analysis.tex
small changes in the document

\newcommand{\bbone}{\mathds{1}}
\DeclareMathOperator*{\diam}{diam}
\DeclareMathOperator*{\spann}{span}
\DeclareMathOperator*{\init}{init}
\DeclareMathOperator*{\ext}{ext}
\newcommand{\udlrarrow}{\mathrlap{\,{\scriptscriptstyle \updownarrow}}{\scriptscriptstyle \leftrightarrow}}

@@

Denote the $2 \times 2$ block-matrix with $M$. We use MINRES with initial guess $x_0$
equal to the right-hand side $b$ of the system, and continue until the relative residual is `small':
\[
  \frac{\|Mx_k - b\|_2}{\|x_k\|_2} \leq 10^{-5}.
\]
This gives us a discrete solution $(\mu_\delta, u_\delta)$ for which we can measure
\[
  E(t) := t \mapsto \|\gamma_t(u_\delta - u)\|_{L_2(\Omega)}.
\]
We will use this as a proxy for the error $\|u - u_\delta\|_X$ which is difficult to measure.

Given $X_\delta$, we distinguish its dimension $\dim X_\delta$ and $\# X_\delta$,
being the total number of \emph{degrees of freedom}, i.e., the dimension of $X_\delta$
minus the number of Dirichlet double nodes (double nodes for which the vertex of
$\lambda_{\text{space}}$ is not on the domain boundary).

\section*{Convergence rate}
The code right now seems to solve the problem, as can be seen by Figure~\ref{fig:time-slice}:
Increasing $L$ (thereby increasing $\# X_\delta$, yields an error reduction.
We discern a pre-asymptotic regime: the error hardly decreases between levels $L=1$
and $2$. Moreover, errors seem to taper off near the final levels $L$, which may
have something to do with our solver tolerance.

@@

\section*{Computational cost}
Let's measure the time it takes to perform a single application of the system
matrix. We do not take into account the setup cost (like
computing $\Sigma$ and $\Theta$). Figure~\ref{fig:linearity} paints a slightly
demotivating picture, in that the bottom graph is not constant. We do have to
place an asterisk here: for small systems, the produced numbers are averaged over
a small number of iterations, so they change a lot from run to run.
\begin{figure}[h!]
  \includegraphics[width=0.8\linewidth]{error_reduction_linearity.pdf}
  \caption{Top: time (s) per apply; bottom: time (ms) per apply per DoF.}
  \label{fig:linearity}
\end{figure}

\section*{The solver}
As mentioned before, we use MINRES with default parameters (tolerance $10^{-5}$,
initial guess $x_0$ equal to the RHS $b$). In Figure~\ref{fig:minres}, we show
the residual norm $\|M x_k - b\|_2$ as a function of $k$ for a selection of solves.
\begin{figure}[h!]
  \includegraphics[width=0.495\linewidth]{error_reduction_minres_history_semilogy.pdf}
  \includegraphics[width=0.495\linewidth]{error_reduction_minres_history_loglog.pdf}
  \caption{Left: semi-log-y MINRES convergence graph for some solves. Right: Same results in log-log format.}
  \label{fig:minres}
\end{figure}
We see that the number of iterations increases strongly, suggesting that we could
definitely use a good preconditioner.

>>>>

\newcommand{\inp}[2]{\langle #1, #2 \rangle}
\newcommand{\bbone}{\mathds{1}}
\DeclareMathOperator*{\diam}{diam}
\DeclareMathOperator*{\spann}{span}
\DeclareMathOperator*{\init}{init}
\DeclareMathOperator*{\ext}{ext}
\newcommand{\udlrarrow}{\mathrlap{\,{\scriptscriptstyle \updownarrow}}{\scriptscriptstyle \leftrightarrow}}

@@

Denote the $2 \times 2$ block-matrix with $M$. We use MINRES with initial guess $x_0$
equal to the right-hand side $b$ of the system, and continue until the relative residual norm is `small':
\[
  \frac{\|Mx_k - b\|_2}{\|b\|_2} \leq 10^{-5}.
\]
This gives us a discrete solution $(\mu_\delta, u_\delta)$ for which we can measure
\[
  E(t) := t \mapsto \|\gamma_t(u_\delta - u)\|_{L_2(\Omega)}
\]
by computing
\[
  E(t)^2 = \inp{u(t)}{u(t)}_{L_2(\Omega)} + \inp{u_\delta(t)}{u_\delta(t)}_{L_2(\Omega)} - 2 \inp{u(t)}{u_\delta(t)}_{L_2(\Omega)}
\]
where we have the first term analytically, the second number can be computed by
an applicator of the space-time mass matrix, and the final term is approximated
using quadrature of sufficient degree. We will use this as a proxy for the error
$\|u - u_\delta\|_X$, which is difficult to measure.

Given $X_\delta$, we distinguish the dimension $\dim X_\delta$ of the function space
and $\# X_\delta$, being the total number of double nodes in the double tree.

\section*{Convergence rate}
The code right now seems to solve the problem, as can be seen by Figure~\ref{fig:time-slice}:
Increasing $L$ (thereby increasing $\dim X_\delta$, yields an error reduction.
We discern a pre-asymptotic regime: the error hardly decreases between levels $L=1$
and $2$. Moreover, errors seem to taper off near the final levels $L$, which may
have something to do with our solver tolerance.

@@

\section*{Computational cost}
Let's measure the time it takes to perform a single application of the system
matrix. We do not take into account the setup cost (like
computing $\Sigma$ and $\Theta$). Figure~\ref{fig:linearity} paints a slightly
demotivating picture, in that the bottom graph is not constant. We do have to
place an asterisk here: for small systems, the produced numbers are averaged over
a small number of iterations, so they change a lot from run to run.
\begin{figure}[h!]
  \includegraphics[width=0.6\linewidth]{error_reduction_linearity.pdf}
  \caption{Top: time (s) per apply; bottom: time (ms) per apply per DoF.}
  \label{fig:linearity}
\end{figure}

\section*{The solver}
As mentioned before, we use MINRES with default parameters (tolerance $10^{-5}$,
initial guess $x_0$ equal to the RHS $b$). In Figure~\ref{fig:minres}, we show
the \emph{absolute} residual norm $\|M x_k - b\|_2$ as a function of $k$ for a
selection of solves.
\begin{figure}[h!]
  \includegraphics[width=0.495\linewidth]{error_reduction_minres_history_semilogy.pdf}
  \includegraphics[width=0.495\linewidth]{error_reduction_minres_history_loglog.pdf}
  \caption{Left: semi-log-y MINRES convergence graph for some solves. Right: Same results in log-log format.}
  \label{fig:minres}
\end{figure}
We see that the number of iterations increases strongly, suggesting that we could
definitely use a good preconditioner.
